---
title: "Michaelis-Menten Modelling"
author: "Nicole J. Dilernia"
date: "2023-07-05"
output:
  html_document: default
  pdf_document: default 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

# Set a working directory 
```{r}
setwd("...")
```


1. Import Data and Load Libraries
# Data Frame - import .csv file, check for NAs, and prepare "clean" dataset without NAs
```{r}
# e.g., 
AabrotanoidesINNERrep1 <- read.csv("...")
   str(AabrotanoidesINNERrep1)

anyNA(AabrotanoidesINNERrep1)
sum(is.na(AabrotanoidesINNERrep1))

AabrotanoidesINNERrep1_clean <- na.omit(AabrotanoidesINNERrep1)
  print(AabrotanoidesINNERrep1_clean)
anyNA(AabrotanoidesINNERrep1_clean)
sum(is.na(AabrotanoidesINNERrep1_clean))
```

# Install and load libraries
```{r}
install.packages("ggplot2") # to plot models
library(ggplot2) 

install.packages("drc") # to fit and plot Michaelis-Menten (MM) model
library(drc)  

# https://www.rdocumentation.org/packages/drc/versions/2.5-12/topics/MM 
```

2. Two parameter model using Michaelis-Menten equation ("drc" package)
# Making/Fitting a model for Michaelis-menten (2 parameters)
```{r}
# Fit the Michaelis-Menten model 
MichaelisModel <- drm(VO2~PO2, data = AabrotanoidesINNERrep1_clean, fct = MM.2())
  
# Create a data frame of predicted values (calculating the data points to make the "MM Line")
MMLine <- data.frame(PO2 = seq(0, max(AabrotanoidesINNERrep1_clean$PO2), length.out = 100)) 
MMLine$VO2 <- predict(MichaelisModel, newdata = MMLine) 
  summary(MichaelisModel)

# Calculate SSR (Sum of Squares Residuals)
ssr <- sum(resid(MichaelisModel)^2) 
  print(ssr) 

# Calculate & show the AIC of the model  
AIC(object = MichaelisModel, k = 2) 

```

# Plotting Michaelis-mentin model on the data
```{r}
# Plot the Michaelis-Menten model against the data
ggplot(AabrotanoidesINNERrep1_clean, aes(x=PO2, y=VO2, colour=factor("VO2"))) +
  geom_point(stat="identity", size = 4) + 
  ylim(0, 0.65) +
  xlim(0, 100) +
  geom_line(data = MMLine, aes(x=PO2, y=VO2), size=2)  +
  scale_colour_brewer(palette = "Blues") +
theme(plot.title = element_text(size = 10),
      axis.title = element_text(size = 20),
      axis.text = element_text(colour = "black", size = 30),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "transparent"), 
      plot.background = element_rect(fill = "transparent", colour = NA),
      axis.line = element_line(colour = "black", size = 2, linetype = 1),
      axis.ticks = element_line(colour = "black", size = 2),
      axis.ticks.length = unit(.4, "cm")) +
      labs(title="AabrotanoidesINNERrep1", x="pO2 (% air sat)", y="VO2 (mg O2/hr)") 
ggsave("AabrotanoidesINNERrep1.pdf", width=10, height=7) 
```

# Display RColorBrewer Palette (color-blind friendly palette)
```{r}
library("RColorBrewer")       
display.brewer.all(colorblindFriendly = TRUE)      # Show all color palettes to pick one
```




